Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1].}

### 2018-11-29 paper/paper.tex
Even more things

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we design an
    algorithm which instead uses degree to direct the proportion of search effort spent in different
    subproblems. This increases performance by two orders of magnitude on satisfiable instances,
    whilst not affecting performance on unsatisfiable instances. This algorithm can also be
    parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
    instances, we get a further speedup of over thirty from thirty-six cores, or ninety five from
    ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
    optimisation problems, by using it to improve two maximum common induced subgraph algorithms.
\end{abstract}

@@

Firstly, we define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we
select a vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') =
\frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely
to pick between vertices of equal degree, is twice as likely to pick a vertex of degree $d$ as one
of degree $d - 1$, and is over a thousand times more likely to pick a vertex of degree $d$ than
degree $d - 10$.

@@

To further test the scalability of this technique, we also implemented a distributed-memory
parallelism layer on top of the threaded layer, using MPI. In contrast to the huge difficulties of
implementing work-stealing in a distributed memory setting, this required only the addition of two
MPI calls: an ``all gather'' operation to communicate nogoods, and a ``gather'' to collect and
combine the results of each host. Termination, meanwhile, was handled by posting an empty nogood
(and so termination can only occur on a restart).

\Cref{figure:parallel} also shows the results of these experiments, using five and ten hosts.
Because each host has two sockets, the five host results use ten MPI ranks, each with eighteen
threads, and the ten host results use twenty MPI ranks. Because the supercomputing service we use is
not designed for huge numbers of very short problems, we ran only instances whose sequential runtime
was at least one second, and for the sake of plotting results, treat skipped instances as taking one
second. Due to the job launcher used, it is also not possible to accurately measure ``total''
runtime including startup costs, and so instead we report the runtime of the rank zero
process---this figure is somewhat optimistic for very easy instances. With these caveats in mind, we
see that our results continue to improve as the host count increases, and we come very close to the
magic goal of an overall speedup of one hundred.

>>>>

\begin{abstract}
    The current state of the art in subgraph isomorphism solving involves using degree as a
    value-ordering heuristic to direct backtracking search. Such a search makes a heavy commitment
    to the first branching choice, which is often incorrect. To mitigate this, we introduce and
    evaluate a new approach, which we call ``solution-biased search''. By combining a
    slightly-random value-ordering heuristic, rapid restarts, and nogood recording, we design an
    algorithm which instead uses degree to direct the proportion of search effort spent in different
    subproblems. This increases performance by two orders of magnitude on satisfiable instances,
    whilst not affecting performance on unsatisfiable instances. This algorithm can also be
    parallelised in a very simple but effective way: across both satisfiable and unsatisfiable
    instances, we get a further speedup of over thirty from thirty-six cores, or ninety-five from
    ten distributed-memory hosts. Finally, we show that solution-biased search is also suitable for
    optimisation problems, by using it to improve two maximum common induced subgraph algorithms.
\end{abstract}

@@

We first define a new \textbf{Biased} value-ordering heuristic, as follows. When branching, we
select a vertex $v'$ from the chosen domain $D_v$ with probability \[ p(v') =
\frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This heuristic is now equally likely
to pick between vertices of equal degree, is twice as likely to pick a vertex of degree $d$ as one
of degree $d - 1$, and is over a thousand times more likely to pick a vertex of degree $d$ than
degree $d - 10$.

@@

To further test the scalability of this technique, we also used MPI to implement a distributed-memory
parallelism layer on top of the threaded layer. In contrast to the huge difficulties of
implementing work-stealing in a distributed memory setting, this required only the addition of two
MPI calls: an ``all gather'' operation to communicate nogoods, and a ``gather'' to collect and
combine the results of each host. Termination, meanwhile, was handled by posting an empty nogood
(and so termination can only occur on a restart).

\Cref{figure:parallel} also shows the results of these experiments, using five, ten, and twenty
hosts.  Because each host has two sockets, the five host results use ten MPI ranks, each with
eighteen threads, and the ten and twenty host results use twenty and forty MPI ranks respectively.
The supercomputing service we use is not designed for huge numbers of very short problems, and so we
ran only instances whose sequential runtime was at least one second; for the sake of plotting
results, we treat skipped instances as taking one second. Due to the job launcher used, it is also
not possible to accurately measure ``total'' runtime including startup costs, and so instead we
report the runtime of the rank zero process---this figure is somewhat optimistic for very easy
instances. With these caveats in mind, when seeing how long a timeout is needed to solve any 14,370
instances, we get aggregate speedups of 57.1 from five hosts and 95.5 from ten hosts over a
sequential baseline; using twenty hosts is slightly slower, due to increased overheads. However,
when seeing how long a timeout is needed to solve 14,424 instances, distributed parallelism gives us
further aggregate speedups of 2.2, 3.5, and 7.9 from five, ten, and twenty hosts, over the single
machine parallel baseline.
